---
title: "GAMs with daylength"
author: "Philippine Chambault"
date: "2023-06-27"
output:
  bookdown::html_document2:
    number_sections: yes
    code_folding: show
    df_print: default
    fig_caption: yes
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: no
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo      = TRUE,
  fig.align = "center",
  message = FALSE,
  warning = FALSE,
  # cache   = FALSE,
  out.width = "100%"
)
```


```{r, global_options, include=FALSE}
library(mgcv)
library(ggcorrplot)
library(ggplot2)
library(tidyquant)
library(gridExtra)
library(tidyverse)
library(dplyr)
library(viridis)
library(DT)
library(data.table)
library(GGally)
```


* Context: obtained different results with the GAMs late June when checking the code, especially some gaps in the predicted values for depth between 110 and 144 m.
* Objective: investigate why such weird residuals.
* changes made compared to previous version (before June 2023):
- retained only months from Jun to Dec for LR dataset
- removed from HR dataset id 27262 that was tracked for less than 1 month
- recalculated mean_dep for HR dataset: previously was mean(mean_dep) instead of mean(max_dep)
- obtained higher deviance but weird residuals




```{r, echo=F}
# load low resolution dataset
# keep ids tracked for at least 1 month
#---------------------------------------
lr <- readRDS("./RDATA/0c.daylength_depth_LR_19ids_tzCorrected.RDS") %>%
  mutate(month = format(date, "%b")) 

lr %>%
  group_by(id) %>%
  summarise(start    = first(date),
            end      = last(date),
            duration = round(difftime(last(date), first(date), units="days")),
            ndives   = n()) 
lr = lr %>% 
  filter(id != "37235",id != "22849b", id != "22850b", # remove HR tags
         id != "27262", id != "27262b", id != "93100", # remove HR tags
         id != "22849", id != "22853",
         id != "37227", id != "37228", id != "27261",  # according to geom_smooth
         id != "7617", id != "7618") %>%
  select(c(id,date,mean_dep,max_dep,daylength,month,sunrise,sunset)) 
         
# load HR dataset
#-----------------
hr <- readRDS("./RDATA/0b.daylength_depth_HR_5ids_tzCorrected.RDS") %>%
  mutate(month = format(date, "%b")) %>%
  filter(month != "Jan")  %>%
  select(c(id,date,mean_dep,max_dep,daylength,month,sunrise,sunset)) %>%
  filter(id != "27262")   # remove this id because tracked over only one month
hr %>%
  group_by(id) %>%
  summarise(start    = first(date),
            end      = last(date),
            duration = round(difftime(last(date), first(date), units="days")),
            ndives   = n()) 

dat = rbind(lr, hr)
length(unique(dat$id))        
dat = dat[!is.na(dat$daylength),] 
```






# Depth vs daylength from raw data {.tabset}

## mean_dep
```{r, echo=FALSE, fig.pos="placeHere", fig.height=4, fig.cap="Relationship between daily mean depth (from max_dep/dive) and daylengtn."}
ggplot(dat, aes(y=mean_dep, x=daylength)) +
  geom_point(size = 0.5) +
  geom_smooth() +
  facet_wrap(~id, ncol=5) +
  theme_tq()
```

## max_dep
```{r, echo=FALSE, fig.pos="placeHere", fig.height=4, fig.cap="Relationship between daily max depth and daylengtn."}
ggplot(dat, aes(y=max_dep, x=daylength)) +
  geom_point(size = 0.5) +
  geom_smooth() +
  facet_wrap(~id, ncol=5) +
  theme_tq()
```





# Pairwise colinearity
```{r, echo=F}
# Create data 
data <- dat %>% dplyr::select(-c(id,date,month,sunrise,sunset)) 

# Function to return points and geom_smooth
# allow for the method to be changed
my_fn <- function(data, mapping, method="loess", ...){
  p <- ggplot(data = data, mapping = mapping) + 
    geom_point(size = 0.3) + 
    geom_smooth(method=method, ..., size = 0.4)
  p
}
```

```{r, echo=FALSE, fig.pos="placeHere", fig.height=4, fig.cap="Pairwise correlations between variables."}
ggpairs(data, 
        lower = list(continuous = my_fn), 
        upper = list(continuous = wrap("cor", size = 2))) +
  theme_tq()
```






# Simple GAM with mean_dep {.tabset}
## Output
* Good deviance even though slightly lower than before.
```{r, echo=F}
dat$id = as.factor(dat$id)
m = gam(mean_dep ~ s(daylength, k=5) + 
          s(id, bs = 're') +            # random intercept
          s(daylength, id, bs = 're'),  # random slope
        data = dat, method="REML")
summary(m) 
```

## Smooth curves
* Good fit
```{r, echo=FALSE, fig.pos="placeHere", fig.height=5.5, fig.cap="GAM output using mean_dep, ID on slope and intercept."}
plot(m,pages=1, shade=T)
```

## Residuals
* Good fit.
```{r, echo=FALSE, fig.pos="placeHere", fig.height=5.5, fig.cap="Diagnostic plots using mean_dep."}
par(mfrow=c(2,2),mar=c(6,5,6,3))
qq.gam(m,cex=1,pch=20)
hist(residuals(m), xlab="Residuals", main="Histogram of residuals")
observed.y <- napredict(m$na.action, m$y)
plot(fitted(m), observed.y, pch=20, xlab = "Fitted Values",
     ylab = "Response", main = "Response vs. Fitted Values")
acf(resid(m),lag.max=200, main="Autocorrelation")
```

## Individual and population curves
```{r, echo=F}
# predictions at individual level
#----------------------------------
ind_pred_inter <- expand.grid(daylength = as.numeric(3:max(dat$daylength,na.rm=T)),
                              id = unique(dat$id))

# add group_to_compare column
ind_pred <- ind_pred_inter %>%
  # set as data.table
  setDT(.) %>% 
  # sort by group_to_compare, time and id
  .[order(daylength, id),] %>%
  # add individual prediction
  .[, fit_ind := predict.gam(m,
                             .SD,
                             type = "response")] %>%
  # add individual SE
  .[, se_ind := predict.gam(m,
                            .SD,
                            se.fit=TRUE,
                            type = "response")$se.fit] %>%
  # sort by group_to_compare, time and id
  .[order(daylength, id),] %>%
  # trick to avoid calling twice this object in the console for display
  .[]


#----------------------------------
# predictions at population level
#----------------------------------
pop_pred <- setDT(
  expand.grid(
    daylength = as.numeric(3:max(dat$daylength,na.rm=T)),
    id = "population_level",
    stringsAsFactors = FALSE
  )
) %>%
  # sort by group_to_compare, time and id
  .[order(daylength, id),] %>%
  # retrieve population prediction
  .[, fit_pop := predict.gam(m,
                             # select the best individual by group_to_compare
                             ind_pred[, .SD[id == "22849b"]],#first(id)
                             type = "response",
                             exclude = c("s(id)",
                                         "s(daylength,id)"))] %>%
  # add standard error at the population level
  .[, fit_pop_se := predict.gam(
    m,
    # select the first individual by group_to_compare
    ind_pred[, .SD[id == "22849b"]],
    type = "response",
    exclude = c("s(id)",
                "s(daylength,id)"),
    se.fit = T
  )$se.fit] %>%
  # trick to avoid calling twice this object in the console for display
  .[]
pop_pred = as_tibble(pop_pred)
```

* Similar curve as before
```{r, echo=FALSE, fig.pos="placeHere", fig.height=5.5, fig.cap="Individual and population smooth curves using daily mean_dep."}
ggplot(pop_pred, aes(x = daylength, y = fit_pop)) +
  geom_ribbon(aes(ymin=fit_pop-fit_pop_se,ymax=fit_pop+fit_pop_se),
              alpha=0.3, fill="mediumseagreen") +
  geom_line(colour="mediumseagreen",lwd=1) +
  geom_line(data=ind_pred, aes(daylength, fit_ind, group=id), 
            colour="mediumseagreen", lwd=0.5, alpha=0.4) +
  ylim(0,120) +
  # geom_point(data=dat, aes(y=mean_dep, x=daylength), alpha=0.3, size=0.5) +
  labs(y = "Daily mean depth (m)", x = "Daylength (hours)") +
  theme_tq()
```





# Simple GAM with max_dep {.tabset}
## Output
* Ok deviance.
```{r, echo=F}
m = gam(max_dep ~ s(daylength, k=5) + 
          s(id, bs = 're') +            # random intercept
          s(daylength, id, bs = 're'),  # random slope
        data = dat, method="REML")
summary(m) 
```

## Smooth curves
* Ok fit for the random intercept, except for 2 IDs.
* Better fit with meandep.
```{r, echo=FALSE, fig.pos="placeHere", fig.height=5.5, fig.cap="GAM output using max_dep, ID on slope and intercept."}
plot(m,pages=1, shade=T)
```

## Residuals
* Ok residuals
```{r, echo=FALSE, fig.pos="placeHere", fig.height=5.5, fig.cap="Diagnostic plots using max_dep."}
par(mfrow=c(2,2),mar=c(6,5,6,3))
qq.gam(m,cex=1,pch=20)
hist(residuals(m), xlab="Residuals", main="Histogram of residuals")
observed.y <- napredict(m$na.action, m$y)
plot(fitted(m), observed.y, pch=20, xlab = "Fitted Values",
     ylab = "Response", main = "Response vs. Fitted Values")
acf(resid(m),lag.max=200, main="Autocorrelation")
```


## Individual and population curves
```{r, echo=F}
#----------------------------------
# predictions at individual level
#----------------------------------
ind_pred_inter <- expand.grid(daylength = as.numeric(3:max(dat$daylength,na.rm=T)),
                              id        = unique(dat$id))

# add group_to_compare column
ind_pred <- ind_pred_inter %>%
  # set as data.table
  setDT(.) %>% 
  # sort by group_to_compare, time and id
  .[order(daylength, id),] %>%
  # add individual prediction
  .[, fit_ind := predict.gam(m,
                             .SD,
                             type = "response")] %>%
  # add individual SE
  .[, se_ind := predict.gam(m,
                            .SD,
                            se.fit=TRUE,
                            type = "response")$se.fit] %>%
  # sort by group_to_compare, time and id
  .[order(daylength, id),] %>%
  # trick to avoid calling twice this object in the console for display
  .[]


#----------------------------------
# predictions at population level
#----------------------------------
pop_pred <- setDT(
  expand.grid(
    daylength = as.numeric(3:max(dat$daylength,na.rm=T)),
    id = "population_level",
    stringsAsFactors = FALSE
  )
) %>%
  # sort by group_to_compare, time and id
  .[order(daylength, id),] %>%
  # retrieve population prediction
  .[, fit_pop := predict.gam(m,
                             # select the best individual by group_to_compare
                             ind_pred[, .SD[id == "22849b"]],#first(id)
                             type = "response",
                             exclude = c("s(id)",
                                         "s(daylength,id)"))] %>%
  # add standard error at the population level
  .[, fit_pop_se := predict.gam(
    m,
    # select the first individual by group_to_compare
    ind_pred[, .SD[id == "22849b"]],
    type = "response",
    exclude = c("s(id)",
                "s(daylength,id)"),
    se.fit = T
  )$se.fit] %>%
  # trick to avoid calling twice this object in the console for display
  .[]
pop_pred = as_tibble(pop_pred)
```

* Ok trend but slope less pronounced than the one using mean depth.
```{r, echo=FALSE, fig.pos="placeHere", fig.height=5.5, fig.cap="Individual and population smooth curves using daily max_dep."}
ggplot(pop_pred, aes(x = daylength, y = fit_pop)) +
  geom_ribbon(aes(ymin=fit_pop-fit_pop_se,ymax=fit_pop+fit_pop_se),
              alpha=0.3, fill="mediumseagreen") +
  geom_line(colour="mediumseagreen",lwd=1) +
  geom_line(data=ind_pred, aes(daylength, fit_ind, group=id), 
            colour="mediumseagreen", lwd=0.5, alpha=0.4) +
  ylim(0,400) +
  # geom_point(data=dat, aes(y=mean_dep, x=daylength), alpha=0.3, size=0.5) +
  labs(y = "Daily max depth (m)", x = "Daylength (hours)") +
  theme_tq()
```





# Simple GAM with mean of mean_dep {.tabset}
```{r, echo=F}
hr2 <- readRDS("./RDATA/0b.daylength_depth_HR_5ids_tzCorrected.RDS") %>%
  mutate(month = format(date, "%b")) %>%
  filter(month != "Jan")  %>%
  select(c(id,date,mean_meandep,max_dep,daylength,month,sunrise,sunset)) %>%
  rename(mean_dep = mean_meandep) %>%
  filter(id != "27262")   # remove this id because tracked over only one month
dat = rbind(lr, hr2)    
dat = dat[!is.na(dat$daylength),]
dat$id = as.factor(dat$id)
```

## Output
* Higher deviance than the one using mean of maxdep.
```{r, echo=F}
m = gam(mean_dep ~ s(daylength, k=5) + 
          s(id, bs = 're') +            
          s(daylength, id, bs = 're'),  
        data = dat, method="REML")
summary(m) 
```

## Smooth curves
* Good fit
```{r, echo=FALSE, fig.pos="placeHere", fig.height=5.5, fig.cap="GAM output using mean of mean_dep, ID on slope and intercept."}
plot(m,pages=1, shade=T)
```

## Residuals
* Good fit
```{r, echo=FALSE, fig.pos="placeHere", fig.height=5.5, fig.cap="Diagnostic plots using mean of mean_dep."}
par(mfrow=c(2,2),mar=c(6,5,6,3))
qq.gam(m,cex=1,pch=20)
hist(residuals(m), xlab="Residuals", main="Histogram of residuals")
observed.y <- napredict(m$na.action, m$y)
plot(fitted(m), observed.y, pch=20, xlab = "Fitted Values",
     ylab = "Response", main = "Response vs. Fitted Values")
acf(resid(m),lag.max=200, main="Autocorrelation")
```


## Individual and population curves
```{r, echo=F}
#----------------------------------
# predictions at individual level
#----------------------------------
ind_pred_inter <- expand.grid(daylength = as.numeric(3:max(dat$daylength,na.rm=T)),
                              id = unique(dat$id))

# add group_to_compare column
ind_pred <- ind_pred_inter %>%
  # set as data.table
  setDT(.) %>% 
  # sort by group_to_compare, time and id
  .[order(daylength, id),] %>%
  # add individual prediction
  .[, fit_ind := predict.gam(m,
                             .SD,
                             type = "response")] %>%
  # add individual SE
  .[, se_ind := predict.gam(m,
                            .SD,
                            se.fit=TRUE,
                            type = "response")$se.fit] %>%
  # sort by group_to_compare, time and id
  .[order(daylength, id),] %>%
  # trick to avoid calling twice this object in the console for display
  .[]


#----------------------------------
# predictions at population level
#----------------------------------
pop_pred <- setDT(
  expand.grid(
    daylength = as.numeric(3:max(dat$daylength,na.rm=T)),
    id = "population_level",
    stringsAsFactors = FALSE
  )
) %>%
  # sort by group_to_compare, time and id
  .[order(daylength, id),] %>%
  # retrieve population prediction
  .[, fit_pop := predict.gam(m,
                             # select the best individual by group_to_compare
                             ind_pred[, .SD[id == "22849b"]],#first(id)
                             type = "response",
                             exclude = c("s(id)",
                                         "s(daylength,id)"))] %>%
  # add standard error at the population level
  .[, fit_pop_se := predict.gam(
    m,
    # select the first individual by group_to_compare
    ind_pred[, .SD[id == "22849b"]],
    type = "response",
    exclude = c("s(id)",
                "s(daylength,id)"),
    se.fit = T
  )$se.fit] %>%
  # trick to avoid calling twice this object in the console for display
  .[]
pop_pred = as_tibble(pop_pred)
```

* Ok trend similar to the one using mean depth.
```{r, echo=FALSE, fig.pos="placeHere", fig.height=5.5, fig.cap="Individual and population smooth curves using daily mean of mean_dep."}
ggplot(pop_pred, aes(x = daylength, y = fit_pop)) +
  geom_ribbon(aes(ymin=fit_pop-fit_pop_se,ymax=fit_pop+fit_pop_se),
              alpha=0.3, fill="mediumseagreen") +
  geom_line(colour="mediumseagreen",lwd=1) +
  geom_line(data=ind_pred, aes(daylength, fit_ind, group=id), 
            colour="mediumseagreen", lwd=0.5, alpha=0.4) +
  ylim(0,130) +
  labs(y = "Daily mean of mean_depth (m)", x = "Daylength (hours)") +
  theme_tq()
```