---
title: "Effect of sea temperature on HP diving behaviour"
author: "Philippine Chambault"
date: "2023-06-14"
output:
  bookdown::html_document2:
    number_sections: yes
    code_folding: show
    df_print: default
    fig_caption: yes
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo      = TRUE,
  fig.align = "center",
  message = FALSE,
  warning = FALSE,
  # cache   = FALSE,
  out.width = "100%"
)
```


```{r, global_options, include=FALSE}
library(tidyverse)
library(ggplot2)
library(raster)
library(viridis)
library(maps)
library(gridExtra)
library(tidyquant)
library(marmap)
library(rnaturalearth)
library(rnaturalearthdata)
library(RColorBrewer) 
library(tidync)
library(data.table)
```


* Context: sea temperature data extracted from CMEMS product PHY-030 (resolution: daily, 0.08 decimal degree).
* Objective: investigate the temporal and spatial patterns of water temperature at the porpoises locations to assess any relationship with the porpoises diving behaviour.



```{r, echo=F}
dive <- readRDS("./RDATA/0b.diveSummary_5HP_calib_5m_zoc0_tzCorrected.RDS") %>%
  dplyr::select(id,dive,start_local,end_local,lon,lat,maxdep) %>%
  mutate(month = format(start_local, "%b"),
         year  = format(start_local, "%Y"))
dive$month = factor(dive$month, 
                    levels=c("Jul","Aug","Sep","Oct","Nov","Dec"))

north_map = map_data("world") %>% group_by(group)
shore     = north_map[north_map$region=="Canada" 
                      | north_map$region=="Greenland"
                      | north_map$region=="Iceland",]

temp <- readRDS("./RDATA/7.Env_Data/temp_0-380m_Oct2013-Dec2014_PHY030.RDS")
  # rename(lon = longitude, 
  #        lat = latitude,
  #        depth = temp_dep) %>%
  # mutate(month = format(time, "%b"),
  #        year  = format(time, "%Y"))
gc()
temp$month = factor(temp$month, 
                   levels=c("Jul","Aug","Sep","Oct","Nov","Dec"))

monthly = temp %>%
  group_by(month,year,lon,lat,depth) %>%
  summarise(temp_mean = mean(thetao)) %>%
  ungroup()
```






# Maps of monthly temperatures in 2013 {.tabset}
* Maps of monthly water temperature in West Greenland extracted at 30 different depths (0 to 380 m) in 2013.
* From 55 m and beyond: CMEMS product contains empty cells close to shore likely because they must apply a bathymetric filter (e.g., ocean products cannot estimate water temperature at depth when below the seafloor close to shore). I contacted the CMEMS team today to confirm that (waiting for their reply), but the risk is that most porpoise locations will be in areas not associated with temperature data.

## At surface
```{r, echo=F}
dat = monthly[monthly$year==2013 & monthly$depth < 5,]
dat$month = factor(dat$month, 
                   levels=c("Jul","Aug","Sep","Oct","Nov","Dec"))
```

```{r, echo=FALSE, fig.pos="placeHere", fig.height=4, fig.cap="Monthly maps of surface temperature in West Greenand in 2013. Red dots refer to the porpoises locations close to the temperature depth."}
ggplot(shore, aes(long, lat)) +
  geom_contour_filled(data=dat, 
                      aes(lon,lat,z=temp_mean),
                      breaks=seq(-2,11,by=1.5)) +
  coord_map("azequidistant", xlim=c(-60,-48), ylim=c(58,72)) +
  geom_polygon(aes(group=group), fill="lightgrey",lwd=1) +
  facet_wrap(~month, ncol=3) +
  # geom_point(data=dive[dive$year=="2013" & dive$maxdep<5,],
  #            aes(x=lon, y=lat), size=0.1, stroke = 1, col = "red") +
  labs(x = "", y = "", title = "Surface temperature",
       fill = "(deg C)") +
  theme_tq() +
  theme(panel.spacing = unit(0, "lines"),
        axis.text = element_blank(),
        legend.position = "bottom")
```

## At 10 m
```{r, echo=F}
dat = monthly[monthly$year==2013 &
                monthly$depth < 10 & monthly$depth > 8,]
dat$month = factor(dat$month, 
                   levels=c("Oct","Nov","Dec"))
```
```{r, echo=FALSE, fig.pos="placeHere", fig.height=4, fig.cap="Monthly maps of temperature at 10 m in West Greenand in 2013."}
ggplot(shore, aes(long, lat)) +
  geom_contour_filled(data=dat, 
                      aes(lon,lat,z=temp_mean),
                      breaks=seq(-2,11,by=1.5)) +
  coord_map("azequidistant", xlim=c(-60,-48), ylim=c(58,72)) +
  geom_polygon(aes(group=group), fill="lightgrey",lwd=1) +
  facet_wrap(~month, ncol=3) +
  geom_point(data=dive[dive$year=="2013" & dive$maxdep<10 & dive$maxdep>9,],
             aes(x=lon, y=lat), size=0.1, col = "red") +
  labs(x = "", y = "", title = "Temperature at 10 m",
       fill = "(deg C)") +
  theme_tq() +
  theme(panel.spacing = unit(0, "lines"),
        axis.text = element_blank(),
        legend.position = "bottom")
```

## At 30 m
```{r, echo=F}
dat = monthly[monthly$year==2013 &
                monthly$depth < 30 & monthly$depth > 29,]
dat$month = factor(dat$month, 
                   levels=c("Oct","Nov","Dec"))
```
```{r, echo=FALSE, fig.pos="placeHere", fig.height=4, fig.cap="Monthly maps of temperature at 30 m in West Greenand in 2013."}
ggplot(shore, aes(long, lat)) +
  geom_contour_filled(data=dat, 
                      aes(lon,lat,z=temp_mean),
                      breaks=seq(-2,11,by=1.5)) +
  coord_map("azequidistant", xlim=c(-60,-48), ylim=c(58,72)) +
  geom_polygon(aes(group=group), fill="lightgrey",lwd=1) +
  facet_wrap(~month, ncol=3) +
  geom_point(data=dive[dive$year=="2013" & dive$maxdep<31 & dive$maxdep>29,],
             aes(x=lon, y=lat), size=0.1, col = "red") +
  labs(x = "", y = "", title = "Temperature at 20 m",
       fill = "(deg C)") +
  theme_tq() +
  theme(panel.spacing = unit(0, "lines"),
        axis.text = element_blank(),
        legend.position = "bottom")
```

## At 55 m
```{r, echo=F}
dat = monthly[monthly$year==2013 &
                monthly$depth < 56 & monthly$depth > 55,]
dat$month = factor(dat$month, 
                   levels=c("Oct","Nov","Dec"))
```
```{r, echo=FALSE, fig.pos="placeHere", fig.height=4, fig.cap="Monthly maps of temperature at 55 m in West Greenand in 2013."}
ggplot(shore, aes(long, lat)) +
  geom_contour_filled(data=dat, 
                      aes(lon,lat,z=temp_mean),
                      breaks=seq(-2,11,by=1.5)) +
  coord_map("azequidistant", xlim=c(-60,-48), ylim=c(58,72)) +
  geom_polygon(aes(group=group), fill="lightgrey",lwd=1) +
  facet_wrap(~month, ncol=3) +
  geom_point(data=dive[dive$year=="2013" & dive$maxdep<56 & dive$maxdep>54,],
             aes(x=lon, y=lat), size=0.1, col = "red") +
  labs(x = "", y = "", title = "Temperature at 55 m",
       fill = "(deg C)") +
  theme_tq() +
  theme(panel.spacing = unit(0, "lines"),
        axis.text = element_blank(),
        legend.position = "bottom")
```

## At 110 m
```{r, echo=F}
dat = monthly[monthly$year==2013 &
                monthly$depth < 110 & monthly$depth > 109,]
dat$month = factor(dat$month, 
                   levels=c("Oct","Nov","Dec"))
```
```{r, echo=FALSE, fig.pos="placeHere", fig.height=4, fig.cap="Monthly maps of temperature at 110 m in West Greenand in 2013."}
ggplot(shore, aes(long, lat)) +
  geom_contour_filled(data=dat, 
                      aes(lon,lat,z=temp_mean),
                      breaks=seq(-2,11,by=1.5)) +
  coord_map("azequidistant", xlim=c(-60,-48), ylim=c(58,72)) +
  geom_polygon(aes(group=group), fill="lightgrey",lwd=1) +
  facet_wrap(~month, ncol=3) +
  geom_point(data=dive[dive$year=="2013" & dive$maxdep<110 & dive$maxdep>108,],
             aes(x=lon, y=lat), size=0.1, col = "red") +
  labs(x = "", y = "", title = "Temperature at 110 m",
       fill = "(deg C)") +
  theme_tq() +
  theme(panel.spacing = unit(0, "lines"),
        axis.text = element_blank(),
        legend.position = "bottom")
```

## At 155 m
```{r, echo=F}
dat = monthly[monthly$year==2013 &
                monthly$depth < 156 & monthly$depth > 155,]
dat$month = factor(dat$month, 
                   levels=c("Oct","Nov","Dec"))
```
```{r, echo=FALSE, fig.pos="placeHere", fig.height=4, fig.cap="Monthly maps of temperature at 155 m in West Greenand in 2013."}
ggplot(shore, aes(long, lat)) +
  geom_contour_filled(data=dat, 
                      aes(lon,lat,z=temp_mean),
                      breaks=seq(-2,11,by=1.5)) +
  coord_map("azequidistant", xlim=c(-60,-48), ylim=c(58,72)) +
  geom_polygon(aes(group=group), fill="lightgrey",lwd=1) +
  facet_wrap(~month, ncol=3) +
  geom_point(data=dive[dive$year=="2013" & dive$maxdep<156 & dive$maxdep>154,],
             aes(x=lon, y=lat), size=0.1, col = "red") +
  labs(x = "", y = "", title = "Temperature at 155 m",
       fill = "(deg C)") +
  theme_tq() +
  theme(panel.spacing = unit(0, "lines"),
        axis.text = element_blank(),
        legend.position = "bottom")
```


## At 222 m
```{r, echo=F}
dat = monthly[monthly$year==2013 &
                monthly$depth < 223 & monthly$depth > 222,]
dat$month = factor(dat$month, 
                   levels=c("Oct","Nov","Dec"))
```
```{r, echo=FALSE, fig.pos="placeHere", fig.height=4, fig.cap="Monthly maps of temperature at 222 m in West Greenand in 2013."}
ggplot(shore, aes(long, lat)) +
  geom_contour_filled(data=dat, 
                      aes(lon,lat,z=temp_mean),
                      breaks=seq(-2,11,by=1.5)) +
  coord_map("azequidistant", xlim=c(-60,-48), ylim=c(58,72)) +
  geom_polygon(aes(group=group), fill="lightgrey",lwd=1) +
  facet_wrap(~month, ncol=3) +
  geom_point(data=dive[dive$year=="2013" & dive$maxdep<223 & dive$maxdep>221,],
             aes(x=lon, y=lat), size=0.1, col = "red") +
  labs(x = "", y = "", title = "Temperature at 222 m",
       fill = "(deg C)") +
  theme_tq() +
  theme(panel.spacing = unit(0, "lines"),
        axis.text = element_blank(),
        legend.position = "bottom")
```

## At 266 m
```{r, echo=F}
dat = monthly[monthly$year==2013 &
                monthly$depth < 267 & monthly$depth > 226,]
dat$month = factor(dat$month, 
                   levels=c("Oct","Nov","Dec"))
```
```{r, echo=FALSE, fig.pos="placeHere", fig.height=4, fig.cap="Monthly maps of temperature at 266 m in West Greenand in 2013."}
ggplot(shore, aes(long, lat)) +
  geom_contour_filled(data=dat, 
                      aes(lon,lat,z=temp_mean),
                      breaks=seq(-2,11,by=1.5)) +
  coord_map("azequidistant", xlim=c(-60,-48), ylim=c(58,72)) +
  geom_polygon(aes(group=group), fill="lightgrey",lwd=1) +
  facet_wrap(~month, ncol=3) +
  geom_point(data=dive[dive$year=="2013" & dive$maxdep<267 & dive$maxdep>265,],
             aes(x=lon, y=lat), size=0.1, col = "red") +
  labs(x = "", y = "", title = "Temperature at 266 m",
       fill = "(deg C)") +
  theme_tq() +
  theme(panel.spacing = unit(0, "lines"),
        axis.text = element_blank(),
        legend.position = "bottom")
```

## At 318 m
```{r, echo=F}
dat = monthly[monthly$year==2013 &
                monthly$depth < 319 & monthly$depth > 318,]
dat$month = factor(dat$month, 
                   levels=c("Oct","Nov","Dec"))
```
```{r, echo=FALSE, fig.pos="placeHere", fig.height=4, fig.cap="Monthly maps of temperature at 318 m in West Greenand in 2013."}
ggplot(shore, aes(long, lat)) +
  geom_contour_filled(data=dat, 
                      aes(lon,lat,z=temp_mean),
                      breaks=seq(-2,11,by=1.5)) +
  coord_map("azequidistant", xlim=c(-60,-48), ylim=c(58,72)) +
  geom_polygon(aes(group=group), fill="lightgrey",lwd=1) +
  facet_wrap(~month, ncol=3) +
  geom_point(data=dive[dive$year=="2013" & dive$maxdep>300 & dive$maxdep<350,],
             aes(x=lon, y=lat), size=0.1, col = "red") +
  labs(x = "", y = "", title = "Temperature at 318 m",
       fill = "(deg C)") +
  theme_tq() +
  theme(panel.spacing = unit(0, "lines"),
        axis.text = element_blank(),
        legend.position = "bottom")
```

## At 380 m
```{r, echo=F}
dat = monthly[monthly$year==2013 &
                monthly$depth > 350,]
dat$month = factor(dat$month, 
                   levels=c("Oct","Nov","Dec"))
```
```{r, echo=FALSE, fig.pos="placeHere", fig.height=4, fig.cap="Monthly maps of temperature at 380 m in West Greenand in 2013."}
ggplot(shore, aes(long, lat)) +
  geom_contour_filled(data=dat, 
                      aes(lon,lat,z=temp_mean),
                      breaks=seq(-2,11,by=1.5)) +
  coord_map("azequidistant", xlim=c(-60,-48), ylim=c(58,72)) +
  geom_polygon(aes(group=group), fill="lightgrey",lwd=1) +
  facet_wrap(~month, ncol=6) +
  geom_point(data=dive[dive$year=="2013" & dive$maxdep>350,],
             aes(x=lon, y=lat), size=0.1, col = "red") +
  labs(x = "", y = "", title = "Temperature at 380 m",
       fill = "(deg C)") +
  theme_tq() +
  theme(panel.spacing = unit(0, "lines"),
        axis.text = element_blank(),
        legend.position = "bottom")
```








# Maps of monthly temperatures in 2014 {.tabset}
* Maps of monthly water temperature in West Greenland extracted at 30 different depths (0 to 380 m) in 2014.
* Similar to 2013: from 55 m and beyond, CMEMS product contains empty cells close to shore so most porpoise locations associated with deep dives will be in areas with no temperature data in depth.


## At surface
```{r, echo=FALSE, fig.pos="placeHere", fig.height=4, fig.cap="Monthly maps of surface temperature West Greenand in 2014."}
ggplot(shore, aes(long, lat)) +
  geom_contour_filled(data=monthly[monthly$year==2014 & monthly$depth < 5,], 
                      aes(lon,lat,z=temp_mean),
                      breaks=seq(-2,11,by=1.5)) +
  coord_map("azequidistant", xlim=c(-60,-48), ylim=c(58,72)) +
  geom_polygon(aes(group=group), fill="lightgrey",lwd=1) +
  facet_wrap(~month, ncol=6) +
  # geom_point(data=dive[dive$year=="2014" & dive$maxdep<5,],
  #            aes(x=lon, y=lat), size=0.1, col = "red") +
  labs(x = "", y = "", title = "Surface temperature",
       fill = "(deg C)") +
  theme_tq() +
  theme(panel.spacing = unit(0, "lines"),
        axis.text = element_blank(),
        legend.position = "bottom")
```

## At 10 m
```{r, echo=FALSE, fig.pos="placeHere", fig.height=4, fig.cap="Monthly maps of temperature at 10 m West Greenand in 2014."}
ggplot(shore, aes(long, lat)) +
  geom_contour_filled(data=monthly[monthly$year==2014 & 
                                     monthly$depth < 10 & monthly$depth > 8,], 
                      aes(lon,lat,z=temp_mean),
                      breaks=seq(-2,11,by=1.5)) +
  coord_map("azequidistant", xlim=c(-60,-48), ylim=c(58,72)) +
  geom_polygon(aes(group=group), fill="lightgrey",lwd=1) +
  facet_wrap(~month, ncol=6) +
  geom_point(data=dive[dive$year=="2014" & dive$maxdep<10 & dive$maxdep>9,],
             aes(x=lon, y=lat), size=0.1, col = "red") +
  labs(x = "", y = "", title = "Temperature at 10 m",
       fill = "(deg C)") +
  theme_tq() +
  theme(panel.spacing = unit(0, "lines"),
        axis.text = element_blank(),
        legend.position = "bottom")
```

## At 30 m
```{r, echo=FALSE, fig.pos="placeHere", fig.height=4, fig.cap="Monthly maps of temperature at 30 m West Greenand in 2014."}
ggplot(shore, aes(long, lat)) +
  geom_contour_filled(data=monthly[monthly$year==2014 & 
                                  monthly$depth < 30 & monthly$depth > 29,], 
                      aes(lon,lat,z=temp_mean),
                      breaks=seq(-2,11,by=1.5)) +
  coord_map("azequidistant", xlim=c(-60,-48), ylim=c(58,72)) +
  geom_polygon(aes(group=group), fill="lightgrey",lwd=1) +
  facet_wrap(~month, ncol=6) +
  geom_point(data=dive[dive$year=="2014" & dive$maxdep<31 & dive$maxdep>29,],
             aes(x=lon, y=lat), size=0.1, col = "red") +
  labs(x = "", y = "", title = "Temperature at 20 m",
       fill = "(deg C)") +
  theme_tq() +
  theme(panel.spacing = unit(0, "lines"),
        axis.text = element_blank(),
        legend.position = "bottom")
```

## At 55 m
```{r, echo=FALSE, fig.pos="placeHere", fig.height=4, fig.cap="Monthly maps of temperature at 55 m West Greenand in 2014."}
ggplot(shore, aes(long, lat)) +
  geom_contour_filled(data=monthly[monthly$year==2014 & 
                                  monthly$depth < 56 & monthly$depth > 55,], 
                      aes(lon,lat,z=temp_mean),
                      breaks=seq(-2,11,by=1.5)) +
  coord_map("azequidistant", xlim=c(-60,-48), ylim=c(58,72)) +
  geom_polygon(aes(group=group), fill="lightgrey",lwd=1) +
  facet_wrap(~month, ncol=6) +
  geom_point(data=dive[dive$year=="2014" & dive$maxdep<56 & dive$maxdep>54,],
             aes(x=lon, y=lat), size=0.1, col = "red") +
  labs(x = "", y = "", title = "Temperature at 55 m",
       fill = "(deg C)") +
  theme_tq() +
  theme(panel.spacing = unit(0, "lines"),
        axis.text = element_blank(),
        legend.position = "bottom")
```

## At 110 m
```{r, echo=FALSE, fig.pos="placeHere", fig.height=4, fig.cap="Monthly maps of temperature at 110 m West Greenand in 2014."}
ggplot(shore, aes(long, lat)) +
  geom_contour_filled(data=monthly[monthly$year==2014 & 
                                     monthly$depth < 110 & monthly$depth > 109,], 
                      aes(lon,lat,z=temp_mean),
                      breaks=seq(-2,11,by=1.5)) +
  coord_map("azequidistant", xlim=c(-60,-48), ylim=c(58,72)) +
  geom_polygon(aes(group=group), fill="lightgrey",lwd=1) +
  facet_wrap(~month, ncol=6) +
  geom_point(data=dive[dive$year=="2014" & dive$maxdep<110 & dive$maxdep>108,],
             aes(x=lon, y=lat), size=0.1, col = "red") +
  labs(x = "", y = "", title = "Temperature at 110 m",
       fill = "(deg C)") +
  theme_tq() +
  theme(panel.spacing = unit(0, "lines"),
        axis.text = element_blank(),
        legend.position = "bottom")
```

## At 155 m
```{r, echo=FALSE, fig.pos="placeHere", fig.height=4, fig.cap="Monthly maps of temperature at 155 m West Greenand in 2014."}
ggplot(shore, aes(long, lat)) +
  geom_contour_filled(data=monthly[monthly$year==2014 & 
                                     monthly$depth < 156 & monthly$depth > 155,], 
                      aes(lon,lat,z=temp_mean),
                      breaks=seq(-2,11,by=1.5)) +
  coord_map("azequidistant", xlim=c(-60,-48), ylim=c(58,72)) +
  geom_polygon(aes(group=group), fill="lightgrey",lwd=1) +
  facet_wrap(~month, ncol=6) +
  geom_point(data=dive[dive$year=="2014" & dive$maxdep<156 & dive$maxdep>154,],
             aes(x=lon, y=lat), size=0.1, col = "red") +
  labs(x = "", y = "", title = "Temperature at 155 m",
       fill = "(deg C)") +
  theme_tq() +
  theme(panel.spacing = unit(0, "lines"),
        axis.text = element_blank(),
        legend.position = "bottom")
```


## At 222 m
```{r, echo=FALSE, fig.pos="placeHere", fig.height=4, fig.cap="Monthly maps of temperature at 222 m West Greenand in 2014."}
ggplot(shore, aes(long, lat)) +
  geom_contour_filled(data=monthly[monthly$year==2014 & 
                                  monthly$depth < 223 & monthly$depth > 222,], 
                      aes(lon,lat,z=temp_mean),
                      breaks=seq(-2,11,by=1.5)) +
  coord_map("azequidistant", xlim=c(-60,-48), ylim=c(58,72)) +
  geom_polygon(aes(group=group), fill="lightgrey",lwd=1) +
  facet_wrap(~month, ncol=6) +
  geom_point(data=dive[dive$year=="2014" & dive$maxdep<223 & dive$maxdep>221,],
             aes(x=lon, y=lat), size=0.1, col = "red") +
  labs(x = "", y = "", title = "Temperature at 222 m",
       fill = "(deg C)") +
  theme_tq() +
  theme(panel.spacing = unit(0, "lines"),
        axis.text = element_blank(),
        legend.position = "bottom")
```

## At 266 m
```{r, echo=FALSE, fig.pos="placeHere", fig.height=4, fig.cap="Monthly maps of temperature at 266 m West Greenand in 2014."}
ggplot(shore, aes(long, lat)) +
  geom_contour_filled(data=monthly[monthly$year==2014 & 
                                  monthly$depth < 267 & monthly$depth > 266,], 
                      aes(lon,lat,z=temp_mean),
                      breaks=seq(-2,11,by=1.5)) +
  coord_map("azequidistant", xlim=c(-60,-48), ylim=c(58,72)) +
  geom_polygon(aes(group=group), fill="lightgrey",lwd=1) +
  facet_wrap(~month, ncol=6) +
  geom_point(data=dive[dive$year=="2014" & dive$maxdep<267 & dive$maxdep>265,],
             aes(x=lon, y=lat), size=0.1, col = "red") +
  labs(x = "", y = "", title = "Temperature at 266 m",
       fill = "(deg C)") +
  theme_tq() +
  theme(panel.spacing = unit(0, "lines"),
        axis.text = element_blank(),
        legend.position = "bottom")
```

## At 318 m
```{r, echo=FALSE, fig.pos="placeHere", fig.height=4, fig.cap="Monthly maps of temperature at 318 m West Greenand in 2014."}
ggplot(shore, aes(long, lat)) +
  geom_contour_filled(data=monthly[monthly$year==2014 & 
                                     monthly$depth < 319 & monthly$depth > 318,], 
                      aes(lon,lat,z=temp_mean),
                      breaks=seq(-2,11,by=1.5)) +
  coord_map("azequidistant", xlim=c(-60,-48), ylim=c(58,72)) +
  geom_polygon(aes(group=group), fill="lightgrey",lwd=1) +
  facet_wrap(~month, ncol=6) +
  geom_point(data=dive[dive$year=="2014" & dive$maxdep>300 & dive$maxdep<350,],
             aes(x=lon, y=lat), size=0.1, col = "red") +
  labs(x = "", y = "", title = "Temperature at 318 m",
       fill = "(deg C)") +
  theme_tq() +
  theme(panel.spacing = unit(0, "lines"),
        axis.text = element_blank(),
        legend.position = "bottom")
```

## At 380 m
```{r, echo=FALSE, fig.pos="placeHere", fig.height=4, fig.cap="Monthly maps of temperature at 380 m West Greenand in 2014."}
ggplot(shore, aes(long, lat)) +
  geom_contour_filled(data=monthly[monthly$year==2014 & 
                                  monthly$depth < 381 & monthly$depth > 380,], 
                      aes(lon,lat,z=temp_mean),
                      breaks=seq(-2,11,by=1.5)) +
  coord_map("azequidistant", xlim=c(-60,-48), ylim=c(58,72)) +
  geom_polygon(aes(group=group), fill="lightgrey",lwd=1) +
  facet_wrap(~month, ncol=6) +
  geom_point(data=dive[dive$year=="2014" & dive$maxdep>350,],
             aes(x=lon, y=lat), size=0.1, col = "red") +
  labs(x = "", y = "", title = "Temperature at 380 m",
       fill = "(deg C)") +
  theme_tq() +
  theme(panel.spacing = unit(0, "lines"),
        axis.text = element_blank(),
        legend.position = "bottom")
```









# Temperature profiles at HP locations {.tabset}
* At each porpoise location, the water temperature associated with the closest depth to the animal's max depth was extracted. 
* The difference between the temperature depth from CMEMS and the porpoise max depth was on average 0.13 m, but went up to 25 m.
* Also, due to a lack of data in shallow waters close to shore, some temperature values associated with deep dives might not be accurate because far away from the true porpoise location.

```{r, echo=F}
temp_atMaxdep <- readRDS("./RDATA/7.Env_Data/temp_at_IndMaxdep_5HP.RDS")
temp_atMaxdep$month = factor(temp_atMaxdep$month, 
                             levels=c("Jul","Aug","Sep","Oct","Nov","Dec"))
```

## Across months (1 smooth curve/id)
* There is a reverse thermocline occurring from October (temperature increases with depth).
* The 5 porpoises show similar trends.

```{r, echo=FALSE, fig.pos="placeHere", fig.height=4, fig.cap="Water temperature profiles extracted at the closest depth and location of the HP."}
ggplot(temp_atMaxdep, aes(y=thetao, x=maxdep)) +
  geom_point(colour="black", size = 1, stroke = 0, alpha=0.3) +
  geom_smooth(method = "gam", aes(colour=id), linewidth=0.6) +
  scale_color_brewer(palette = "Dark2") +
  coord_flip() +
  scale_x_continuous(trans = "reverse", limits = c(380, 0)) +
  labs(y = "Water temperature (deg C)", x = "Depth (m)") +
  facet_wrap(~month, ncol = 6) +
  theme_tq() +
  theme(legend.position = "bottom",
        panel.spacing.x = unit(0.0, "lines"))
```


## Across months (1 smooth curve for all)
* Typical temperature profile in summer (July to Sep) with decreasing temperature in deeper layer. It's hard to distinguish the mixed layer depth in July and August, more visible in Sep (up to 25-30 m).

```{r, echo=FALSE, fig.pos="placeHere", fig.height=4, fig.cap="Water temperature profiles extracted at the closest depth and location of the HP."}
ggplot(temp_atMaxdep, aes(y=thetao, x=maxdep)) +
  geom_point(colour="black", size = 1, stroke = 0, alpha=0.3) +
  geom_smooth(method = "gam", colour="brown2", linewidth=0.6) +
  scale_color_brewer(palette = "Dark2") +
  coord_flip() +
  scale_x_continuous(trans = "reverse", limits = c(380, 0)) +
  labs(y = "Water temperature (deg C)", x = "Depth (m)") +
  facet_wrap(~month, ncol = 6) +
  theme_tq() +
  theme(legend.position = "bottom",
        panel.spacing.x = unit(0.0, "lines"))
```


## Over time
* Water temp between 0-8 deg C, globally colder below 50 m.
* Some cells overlap because several dives per day were included here. Would be worth plotting the same graph with the average daily depth.

```{r, echo=FALSE, fig.pos="placeHere", fig.height=4, fig.cap="Water temperature profiles over time for the 5 HP."}
ggplot(temp_atMaxdep, aes(y=maxdep, x=start_local)) +
  geom_point(aes(colour=thetao), size = 0.5, stroke = 0) +
  scale_y_continuous(trans = "reverse", limits = c(380, 0)) +
  scale_colour_continuous(type = "viridis") +
  labs(x = "", y = "Depth (m)", 
       colour="Water \ntemperature \n(deg C)") +
  facet_wrap(~id, ncol = 5, scales = "free_x") +
  theme_tq() +
  theme(axis.text.x = element_text(angle = 90),
        panel.spacing.x = unit(0.0, "lines")) 
```



# Concusions
* Interesting analysis suggesting an inversion of the thermocline in October, which might coincide with a shift in the diving behaviour due to a shift in the vertical distribution of the prey.
* Unfortunately, the lack of temperature data in coastal waters where most of the porpoise locations are, prevents investigating the potential link between the porpoises diving behaviour and water temperature.
